### 2024-09-05 _tex/paper.tex
Built site for gh-pages

% Options for packages loaded elsewhere
\PassOptionsToPackage{unicode}{hyperref}
\PassOptionsToPackage{hyphens}{url}
\PassOptionsToPackage{dvipsnames,svgnames,x11names}{xcolor}
%
\documentclass[
]{agujournal2019}

\usepackage{amsmath,amssymb}
\usepackage{iftex}
\ifPDFTeX
  \usepackage[T1]{fontenc}
  \usepackage[utf8]{inputenc}
  \usepackage{textcomp} % provide euro and other symbols
\else % if luatex or xetex
  \usepackage{unicode-math}
  \defaultfontfeatures{Scale=MatchLowercase}
  \defaultfontfeatures[\rmfamily]{Ligatures=TeX,Scale=1}
\fi
\usepackage{lmodern}
\ifPDFTeX\else  
    % xetex/luatex font selection
\fi
% Use upquote if available, for straight quotes in verbatim environments
\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
\IfFileExists{microtype.sty}{% use microtype if available
  \usepackage[]{microtype}
  \UseMicrotypeSet[protrusion]{basicmath} % disable protrusion for tt fonts
}{}
\makeatletter
\@ifundefined{KOMAClassName}{% if non-KOMA class
  \IfFileExists{parskip.sty}{%
    \usepackage{parskip}
  }{% else
    \setlength{\parindent}{0pt}
    \setlength{\parskip}{6pt plus 2pt minus 1pt}}
}{% if KOMA class
  \KOMAoptions{parskip=half}}
\makeatother
\usepackage{xcolor}
\setlength{\emergencystretch}{3em} % prevent overfull lines
\setcounter{secnumdepth}{5}
% Make \paragraph and \subparagraph free-standing
\makeatletter
\ifx\paragraph\undefined\else
  \let\oldparagraph\paragraph
  \renewcommand{\paragraph}{
    \@ifstar
      \xxxParagraphStar
      \xxxParagraphNoStar
  }
  \newcommand{\xxxParagraphStar}[1]{\oldparagraph*{#1}\mbox{}}
  \newcommand{\xxxParagraphNoStar}[1]{\oldparagraph{#1}\mbox{}}
\fi
\ifx\subparagraph\undefined\else
  \let\oldsubparagraph\subparagraph
  \renewcommand{\subparagraph}{
    \@ifstar
      \xxxSubParagraphStar
      \xxxSubParagraphNoStar
  }
  \newcommand{\xxxSubParagraphStar}[1]{\oldsubparagraph*{#1}\mbox{}}
  \newcommand{\xxxSubParagraphNoStar}[1]{\oldsubparagraph{#1}\mbox{}}
\fi
\makeatother

\usepackage{color}
\usepackage{fancyvrb}
\newcommand{\VerbBar}{|}
\newcommand{\VERB}{\Verb[commandchars=\\\{\}]}
\DefineVerbatimEnvironment{Highlighting}{Verbatim}{commandchars=\\\{\}}
% Add ',fontsize=\small' for more characters per line
\usepackage{framed}
\definecolor{shadecolor}{RGB}{241,243,245}
\newenvironment{Shaded}{\begin{snugshade}}{\end{snugshade}}
\newcommand{\AlertTok}[1]{\textcolor[rgb]{0.68,0.00,0.00}{#1}}
\newcommand{\AnnotationTok}[1]{\textcolor[rgb]{0.37,0.37,0.37}{#1}}
\newcommand{\AttributeTok}[1]{\textcolor[rgb]{0.40,0.45,0.13}{#1}}
\newcommand{\BaseNTok}[1]{\textcolor[rgb]{0.68,0.00,0.00}{#1}}
\newcommand{\BuiltInTok}[1]{\textcolor[rgb]{0.00,0.23,0.31}{#1}}
\newcommand{\CharTok}[1]{\textcolor[rgb]{0.13,0.47,0.30}{#1}}
\newcommand{\CommentTok}[1]{\textcolor[rgb]{0.37,0.37,0.37}{#1}}
\newcommand{\CommentVarTok}[1]{\textcolor[rgb]{0.37,0.37,0.37}{\textit{#1}}}
\newcommand{\ConstantTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{#1}}
\newcommand{\ControlFlowTok}[1]{\textcolor[rgb]{0.00,0.23,0.31}{\textbf{#1}}}
\newcommand{\DataTypeTok}[1]{\textcolor[rgb]{0.68,0.00,0.00}{#1}}
\newcommand{\DecValTok}[1]{\textcolor[rgb]{0.68,0.00,0.00}{#1}}
\newcommand{\DocumentationTok}[1]{\textcolor[rgb]{0.37,0.37,0.37}{\textit{#1}}}
\newcommand{\ErrorTok}[1]{\textcolor[rgb]{0.68,0.00,0.00}{#1}}
\newcommand{\ExtensionTok}[1]{\textcolor[rgb]{0.00,0.23,0.31}{#1}}
\newcommand{\FloatTok}[1]{\textcolor[rgb]{0.68,0.00,0.00}{#1}}
\newcommand{\FunctionTok}[1]{\textcolor[rgb]{0.28,0.35,0.67}{#1}}
\newcommand{\ImportTok}[1]{\textcolor[rgb]{0.00,0.46,0.62}{#1}}
\newcommand{\InformationTok}[1]{\textcolor[rgb]{0.37,0.37,0.37}{#1}}
\newcommand{\KeywordTok}[1]{\textcolor[rgb]{0.00,0.23,0.31}{\textbf{#1}}}
\newcommand{\NormalTok}[1]{\textcolor[rgb]{0.00,0.23,0.31}{#1}}
\newcommand{\OperatorTok}[1]{\textcolor[rgb]{0.37,0.37,0.37}{#1}}
\newcommand{\OtherTok}[1]{\textcolor[rgb]{0.00,0.23,0.31}{#1}}
\newcommand{\PreprocessorTok}[1]{\textcolor[rgb]{0.68,0.00,0.00}{#1}}
\newcommand{\RegionMarkerTok}[1]{\textcolor[rgb]{0.00,0.23,0.31}{#1}}
\newcommand{\SpecialCharTok}[1]{\textcolor[rgb]{0.37,0.37,0.37}{#1}}
\newcommand{\SpecialStringTok}[1]{\textcolor[rgb]{0.13,0.47,0.30}{#1}}
\newcommand{\StringTok}[1]{\textcolor[rgb]{0.13,0.47,0.30}{#1}}
\newcommand{\VariableTok}[1]{\textcolor[rgb]{0.07,0.07,0.07}{#1}}
\newcommand{\VerbatimStringTok}[1]{\textcolor[rgb]{0.13,0.47,0.30}{#1}}
\newcommand{\WarningTok}[1]{\textcolor[rgb]{0.37,0.37,0.37}{\textit{#1}}}

\providecommand{\tightlist}{%
  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}\usepackage{longtable,booktabs,array}
\usepackage{calc} % for calculating minipage widths
% Correct order of tables after \paragraph or \subparagraph
\usepackage{etoolbox}
\makeatletter
\patchcmd\longtable{\par}{\if@noskipsec\mbox{}\fi\par}{}{}
\makeatother
% Allow footnotes in longtable head/foot
\IfFileExists{footnotehyper.sty}{\usepackage{footnotehyper}}{\usepackage{footnote}}
\makesavenoteenv{longtable}
\usepackage{graphicx}
\makeatletter
\def\maxwidth{\ifdim\Gin@nat@width>\linewidth\linewidth\else\Gin@nat@width\fi}
\def\maxheight{\ifdim\Gin@nat@height>\textheight\textheight\else\Gin@nat@height\fi}
\makeatother
% Scale images if necessary, so that they will not overflow the page
% margins by default, and it is still possible to overwrite the defaults
% using explicit options in \includegraphics[width, height, ...]{}
\setkeys{Gin}{width=\maxwidth,height=\maxheight,keepaspectratio}
% Set default figure placement to htbp
\makeatletter
\def\fps@figure{htbp}
\makeatother
% definitions for citeproc citations
\NewDocumentCommand\citeproctext{}{}
\NewDocumentCommand\citeproc{mm}{%
  \begingroup\def\citeproctext{#2}\cite{#1}\endgroup}
\makeatletter
 % allow citations to break across lines
 \let\@cite@ofmt\@firstofone
 % avoid brackets around text for \cite:
 \def\@biblabel#1{}
 \def\@cite#1#2{{#1\if@tempswa , #2\fi}}
\makeatother
\newlength{\cslhangindent}
\setlength{\cslhangindent}{1.5em}
\newlength{\csllabelwidth}
\setlength{\csllabelwidth}{3em}
\newenvironment{CSLReferences}[2] % #1 hanging-indent, #2 entry-spacing
 {\begin{list}{}{%
  \setlength{\itemindent}{0pt}
  \setlength{\leftmargin}{0pt}
  \setlength{\parsep}{0pt}
  % turn on hanging indent if param 1 is 1
  \ifodd #1
   \setlength{\leftmargin}{\cslhangindent}
   \setlength{\itemindent}{-1\cslhangindent}
  \fi
  % set entry spacing
  \setlength{\itemsep}{#2\baselineskip}}}
 {\end{list}}
\usepackage{calc}
\newcommand{\CSLBlock}[1]{\hfill\break\parbox[t]{\linewidth}{\strut\ignorespaces#1\strut}}
\newcommand{\CSLLeftMargin}[1]{\parbox[t]{\csllabelwidth}{\strut#1\strut}}
\newcommand{\CSLRightInline}[1]{\parbox[t]{\linewidth - \csllabelwidth}{\strut#1\strut}}
\newcommand{\CSLIndent}[1]{\hspace{\cslhangindent}#1}

\usepackage{url} %this package should fix any errors with URLs in refs.
\usepackage{lineno}
\usepackage[inline]{trackchanges} %for better track changes. finalnew option will compile document with changes incorporated.
\usepackage{soul}
\linenumbers
\makeatletter
\@ifpackageloaded{caption}{}{\usepackage{caption}}
\AtBeginDocument{%
\ifdefined\contentsname
  \renewcommand*\contentsname{Table of contents}
\else
  \newcommand\contentsname{Table of contents}
\fi
\ifdefined\listfigurename
  \renewcommand*\listfigurename{List of Figures}
\else
  \newcommand\listfigurename{List of Figures}
\fi
\ifdefined\listtablename
  \renewcommand*\listtablename{List of Tables}
\else
  \newcommand\listtablename{List of Tables}
\fi
\ifdefined\figurename
  \renewcommand*\figurename{Figure}
\else
  \newcommand\figurename{Figure}
\fi
\ifdefined\tablename
  \renewcommand*\tablename{Table}
\else
  \newcommand\tablename{Table}
\fi
}
\@ifpackageloaded{float}{}{\usepackage{float}}
\floatstyle{ruled}
\@ifundefined{c@chapter}{\newfloat{codelisting}{h}{lop}}{\newfloat{codelisting}{h}{lop}[chapter]}
\floatname{codelisting}{Listing}
\newcommand*\listoflistings{\listof{codelisting}{List of Listings}}
\makeatother
\makeatletter
\makeatother
\makeatletter
\@ifpackageloaded{caption}{}{\usepackage{caption}}
\@ifpackageloaded{subcaption}{}{\usepackage{subcaption}}
\makeatother

\ifLuaTeX
  \usepackage{selnolig}  % disable illegal ligatures
\fi
\usepackage{bookmark}

\IfFileExists{xurl.sty}{\usepackage{xurl}}{} % add URL line breaks if available
\urlstyle{same} % disable monospaced font for URLs
\hypersetup{
  pdftitle={Evaluating Cloud-Optimized HDF5 for NASA's ICESat-2 Mission},
  pdfauthor={Luis A. Lopez; Andrew P. Barrett; Amy Steiker; Aleksandar Jelenak; Lisa Kaser; Jeffrey E. Lee},
  pdfkeywords={cloud-native, cloud, HDF5, NASA, ICESat-2},
  colorlinks=true,
  linkcolor={blue},
  filecolor={Maroon},
  citecolor={Blue},
  urlcolor={Blue},
  pdfcreator={LaTeX via pandoc}}

\journalname{Geophysical Research Letters}

\draftfalse

\begin{document}
\title{Evaluating Cloud-Optimized HDF5 for NASA's ICESat-2 Mission}

\authors{Luis A. Lopez\affil{1}, Andrew P. Barrett\affil{1}, Amy
Steiker\affil{1}, Aleksandar Jelenak\affil{2}, Lisa
Kaser\affil{1}, Jeffrey E. Lee\affil{3}}
\affiliation{1}{CIRES, National Snow and Ice Data Center, University of
Colorado, Boulder., Boulder, CO, USA}\affiliation{2}{The HDF
Group, Champaign, IL, USA}\affiliation{3}{NASA Goddard Space Flight
Center, NASA / KBR, Greenbelt, MD, USA}

\begin{abstract}
The Hierarchical Data Format (HDF) is a common archival format for
n-dimensional scientific data; it has been utilized to store valuable
information from astrophysics to earth sciences and everything in
between. As flexible and powerful as HDF can be, it comes with big
tradeoffs when it's accessed from remote storage systems, mainly because
the file format and the client I/O libraries were designed for local and
supercomputing workflows. As scientific data and workflows migrate to
the cloud , efficient access to data stored in HDF format is a key
factor that will accelerate or slow down ``science in the cloud'' across
all disciplines. We present an implementation of recently available
features in the HDF5 stack that results in performant access to HDF from
remote cloud storage. This performance is on par with modern
cloud-native formats like Zarr but with the advantage of not having to
reformat data or generate metadata sidecar files (DMR++, Kerchunk). Our
benchmarks also show potential cost-savings for data producers if their
data are processed using cloud-optimized strategies.
\end{abstract}

\section{Problem}\label{problem}

Scientific data from NASA and other agencies are increasingly being
distributed from the commercial cloud. Cloud storage enables large-scale
workflows and should reduce local storage costs. It also allows the use
of scalable on-demand cloud computing resources by individual scientists
and the broader scientific community. However, the majority of this
scientific data is stored in a format that was not designed for the
cloud: The Hierarchical Data format or HDF.

The most recent version of the Hierarchical Data Format is HDF5, a
common archival format for n-dimensional scientific data; it has been
utilized to store valuable information from astrophysics to earth
sciences and everything in between. As flexible and powerful as HDF5 can
be, it comes with big tradeoffs when it's accessed from remote storage
systems.

HDF5 is a complex file format; we can think of it as a file system using
a tree-like structure with multiple data types and native data
structures. Because of this complexity, the most reliable way of
accessing data stored in this format is using the HDF5 C API. Regardless
of access pattern, nearly all tools ultimately rely on the HDF5-C
library and this brings a couple issues that affect the efficiency of
accessing this format over the network:

\begin{center}\rule{0.5\linewidth}{0.5pt}\end{center}

\subsubsection{\texorpdfstring{\textbf{Metadata
fragmentation}}{Metadata fragmentation}}\label{metadata-fragmentation}

When working with large datasets, especially those that include numerous
variables and nested groups, the storage of file-level metadata can
become a challenge. By default, metadata associated with each dataset is
stored in chunks of 4 kilobytes (KB). This chunking mechanism was
originally intended to optimize storage efficiency and access speed on
disks with hardware resources available more than 20 years ago. In
datasets with many variables and/or complex hierarchical structures,
these 4KB chunks can lead to significant fragmentation.

Fragmentation occurs when this metadata is spread out across multiple
non-contiguous chunks within the file. This results in inefficiencies
when accessing or modifying data because compatible libraries need to
read from multiple, scattered locations in the file. Over time, as the
dataset grows and evolves, this fragmentation can compound, leading to
degraded performance and increased storage overhead. In particular,
operations that involve reading or writing metadata, such as opening a
file, checking attributes, or modifying variables, can become slower and
more resource-intensive.

\subsubsection{\texorpdfstring{\textbf{Global API
Lock}}{Global API Lock}}\label{global-api-lock}

Because of the historical complexity of operations with the HDF5
format(The HDF Group, n.d.), there has been a necessity to make the
library thread-safe and similarly to what happens in the Python
language, the simplest mechanism to implement this is to have a global
API lock. This global lock is not as big of an issue when we read data
from local disk but it becomes a major bottleneck when we read data over
the network because each read is sequential and latency in the cloud is
exponentially bigger than local access (MDN, 2024) (Scott, 2020).

\begin{center}\rule{0.5\linewidth}{0.5pt}\end{center}

\begin{figure*}

\centering{

\includegraphics{figures/figure-1.png}

}

\caption{\label{fig-1}shows how reads (Rn) are done in order to access
file metadata, In the first read, R0, the HDF5 library verifies the file
signature from the superblock, subsequent reads, R1, R2,\ldots Rn, read
file metadata, 4kb at the time.}

\end{figure*}%

\subsubsection{\texorpdfstring{\textbf{Background and data
selection}}{Background and data selection}}\label{background-and-data-selection}

As a result of community feedback and ``hack weeks'' organized by NSIDC
and UW eScience Institute in 2023(ICESAT-2 HackWeek, 2023), NSIDC
started the Cloud Optimized Format Investigation (COFI) project to
improve access to HDF5 from the ICESat-2 mission. A spaceborne lidar
that retrieves surface topography of the Earth's ice sheets, land and
oceans (Neumann et al., 2019). Because of its complexity, large size and
importance for cryospheric studies we targeted the ATL03 dataset. ATL03
core data are geolocated photon heights from the ICESat-2 ATLAS
instrument. Each file contains 1003 geophysical variables in 6 data
groups. Although our research was focused on this dataset, most of our
findings are applicable to any dataset stored in HDF5 and NetCDF4.

\section{Methodology}\label{methodology}

We tested access times to original and different configurations of
cloud-optimized HDF5 ATL03 files {[}list files tested{]} stored in AWS
S3 buckets in region us-west-2, the region hosting NASA's Earthdata
Cloud archives. Files were accessed using Python tools commonly used by
Earth scientists: h5py and Xarray(Hoyer \& Hamman, 2017). h5py is a
Python wrapper around the HDF5 C API. xarray\footnote{\texttt{h5py} is a
  dependency of Xarray} is a widely used Python package for working with
n-dimensional data. We also tested access times using h5coro, a python
package optimized for reading HDF5 files from S3 buckets and kerchunk, a
tool that creates an efficient lookup table for file chunks to allow
performant partial reads of files.

The test files were originally cloud optimized by ``repacking'' them,
using a relatively new feature in the HDF5 C API called ``paged
aggregation''. Page aggregation does 2 things: it collects file-level
metadata from datasets and stores it on dedicated metadata blocks in the
file; and it forces the library to write both data and metadata using
these fixed-size pages. Aggregation allows client libraries to read file
metadata with only a few requests and uses the page size used in the
aggregation as the minimal request size, overriding the 1 request per
chunk behavior.

\begin{figure*}

\centering{

\includegraphics{figures/figure-2.png}

}

\caption{\label{fig-2}shows how file-level metadata and data gets
internally packed once we use paged aggregation on a file.}

\end{figure*}%

As we can see in Figure~\ref{fig-2}, when we cloud optimize a file using
paged-aggregation there are some considerations and behavior that we had
to take into account. The first thing to observe is that page
aggregation will --as we mentioned-- consolidate the file-evel metadata
at the front of the file and will add information in the so-called
superblock\footnote{The HDF5 superblock is a crucial component of the
  HDF5 file format, acting as the starting point for accessing all data
  within the file. It stores important metadata such as the version of
  the file format, pointers to the root group, and addresses for
  locating different file components} The next thing to notice is that
page size us uses across the board for metadata and data as of October
2024 and version 1.14 of the HDF5 library, page size cannot dynamically
adjust to the total metadata size.

\begin{figure*}

\centering{

\includegraphics{figures/figure-3.png}

}

\caption{\label{fig-3}shows how file-level metadata and data packing
inside aggregated pages leave unused space that can potentially increase
the file size in a considerable way.}

\end{figure*}%

This one page size for all approach simplifies how the HDF5 API reads
the file (if instructucted) but it also brings unused page space and
chunk over-reads. In the case of the ICESat-2 dataset (ATL03) the data
itself has been partitioned and each granule represents a segment in the
satellite orbit and within the file the most relevant dataset is chunked
using 10,000 items per chunk, with data being float-32 and using a fast
compression value, the resulting chunk size is on average under 40KB,
which is really small for an HTTP request, especially when we have to
read them sequentially. Because of these considerations, we opted for
testing different page sizes, and increase chunk size. The following
table describes the different configurations used in our tests.

\begin{longtable}[]{@{}
  >{\raggedright\arraybackslash}p{(\columnwidth - 12\tabcolsep) * \real{0.1589}}
  >{\raggedright\arraybackslash}p{(\columnwidth - 12\tabcolsep) * \real{0.3444}}
  >{\raggedright\arraybackslash}p{(\columnwidth - 12\tabcolsep) * \real{0.1457}}
  >{\raggedright\arraybackslash}p{(\columnwidth - 12\tabcolsep) * \real{0.0993}}
  >{\raggedright\arraybackslash}p{(\columnwidth - 12\tabcolsep) * \real{0.0728}}
  >{\raggedright\arraybackslash}p{(\columnwidth - 12\tabcolsep) * \real{0.0728}}
  >{\raggedright\arraybackslash}p{(\columnwidth - 12\tabcolsep) * \real{0.1060}}@{}}
\toprule\noalign{}
\begin{minipage}[b]{\linewidth}\raggedright
prefix
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
description
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
\% file size increase
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
\textasciitilde km per chunk
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
shape
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
page size
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
avg\_chunk\_size
\end{minipage} \\
\midrule\noalign{}
\endhead
\bottomrule\noalign{}
\endlastfoot
original & original file from ATL03 v6 (1gb and 7gb) & 0 & 1.5km &
(10000,) & N/A & 35kb \\
original-kerchunk & kerchunk sidecar of the original file & N/A & 1.5km
& (10000,) & N/A & 35kb \\
page-only-4mb & paged-aggregated file with 4mb per page &
\textasciitilde1\% & 1.5km & (10000,) & 4MB & 35kb \\
page-only-8mb & paged-aggregated file with 4mb per pag8 &
\textasciitilde1\% & 1.5km & (10000,) & 8MB & 35kb \\
rechunked-4mb & page-aggregated and bigger chunk sizes &
\textasciitilde1\% & 10km & (100000,) & 4MB & 400kb \\
rechunked-8mb & page-aggregated and bigger chunk sizes &
\textasciitilde1\% & 10km & (100000,) & 8MB & 400kb \\
rechunked-8mb-kerchunk & kerchunk sidecar of the last paged-aggregated
file & N/A & 10km & (10000,) & 8MB & 400kb \\
\end{longtable}

This table represents the different configurations we used for our tests
in 2 file sizes. It's worth noticing that we encountered a few outlier
cases where compression and chunk sizes led page aggregation to an
increase in file size of approximately 10\% which was above the desired
value for NSIDC (5\% max) We tested these files using the most common
libraries to handle HDF5 and 2 different I/O drivers that support remote
access to AWS S3, fsspec and the native S3. The results of our testing
is explained in the next section and the code to reproduce the results
is in the attached notebooks.

\section{Results}\label{results}

\begin{figure*}

\centering{

\includegraphics{figures/figure-5.png}

}

\caption{\label{fig-5}Benchmarks show that cloud optimizing ATL03 files
improved access times at least an order of magnitude when used with
aligned I/O patterns, this is telling the library about the cloud
optimization and page size.}

\end{figure*}%

\section{Recommendations}\label{recommendations}

We have split our recommendations for the ATL03 product into 3 main
categories, creating the files, accessing the files, and future tool
development.

\subsection{Recommended cloud
optimizations}\label{recommended-cloud-optimizations}

Based on our testing we recommend the following cloud optimizations for
creating HDF5 files for the ATL03 product: Create HDF5 files using paged
aggregation by setting HDF5 library parameters:

\begin{enumerate}
\def\labelenumi{\arabic{enumi}.}
\tightlist
\item
  File page strategy: H5F\_FSPACE\_STRATEGY\_PAGE
\item
  File page size: 8000000 If repacking an existing file, h5repack
  contains the code to alter tese variables inside the file
\end{enumerate}

\begin{Shaded}
\begin{Highlighting}[]
 \ExtensionTok{h5repack} \AttributeTok{{-}S}\NormalTok{ PAGE }\AttributeTok{{-}G}\NormalTok{ 8000000 input.h5 output.h5}
\end{Highlighting}
\end{Shaded}

\begin{enumerate}
\def\labelenumi{\arabic{enumi}.}
\setcounter{enumi}{2}
\tightlist
\item
  Avoid using unlimited dimensions when creating variables because the
  HDF5 API cannot support it inside bffered pages and representation of
  these variables is not supported by Kerchunk.
\end{enumerate}

\subsubsection{Reasoning}\label{reasoning}

Based on the variable size of ATL03 it becomes really difficult to
allocate a fixed metadata page, big files contain north of 30MB of
metadata, but the median sized file is below 8MB. If we had adopted user
block we would have caused an increase in the file size and storage cost
of approximate 30\% (reference to our tests). Another consequence of
using a dedicated fixed page for file-level metadata is that metadata
overflow will generate the same impact in access times, the library will
fetch the metadata in one go but the rest will be using the predefined
block size of 4kb.

Paged aggregation is thus the simplest way of cloud optimizing an HDF5
file as the metadata will keep filling dedicated pages until all the
file-level metadata is stored at the front of the file. Chunk sizes
cannot be larger than the page size and when chunk sizes are smaller we
need to take into account how these chunks will fit on a page, in an
ideal scenario all the space will be filled but that is not the case and
we will end up with unused space See~\ref{fig-2}.

\subsection{Recommended access
patterns}\label{recommended-access-patterns}

Placeholder

\subsection{Recommended tooling
development}\label{recommended-tooling-development}

Placeholder

\subsection{Mission implementation}\label{mission-implementation}

ATL03 is a complex science data product containing both segmented (20
meters along-track) and large, variable-rate photon datasets. ATL03 is
created using pipeline-style processing where the science data and
NetCDF-style metadata are written by independent software packages. The
following steps were employed to create cloud-optimized Release 007
ATL03 products, while minimizing increases in file size:

\begin{enumerate}
\def\labelenumi{\arabic{enumi}.}
\tightlist
\item
  Set the ``file space strategy'' to H5F\_FSPACE\_STRATEGY\_PAGE and
  enabled ``free space tracking'' within the HDF5 file creation property
  list.
\item
  Set the ``file space page size'' to 8MiB.
\item
  Changed all ``COMPACT'' dataset storage types to ``CONTIGUOUS''.
\item
  Increased the ``chunk size'' of the photon-rate datasets (from 10,000
  to 100,000 elements), while making sure no ``chunk sizes'' exceeded
  the 8MiB ``file space page size''.
\item
  Introduced a new production step that executes the ``h5repack''
  utility (with no options) to create a ``defragmented'' final product.
\end{enumerate}

\section{Discussion}\label{discussion}

\begin{enumerate}
\def\labelenumi{\arabic{enumi}.}
\tightlist
\item
  Chunking shapes and sizes
\item
  Paged aggregation vs User block
\item
  Side effects on different access patterns, e.g.~Kerchunk
\end{enumerate}

\section{References}\label{references}

\phantomsection\label{refs}
\begin{CSLReferences}{1}{0}
\vspace{1em}

\bibitem[\citeproctext]{ref-Hoyer2017-su}
Hoyer, S., \& Hamman, J. (2017). Xarray: {N-D} labeled arrays and
datasets in python. \emph{J. Open Res. Softw.}, \emph{5}(1), 10.

\bibitem[\citeproctext]{ref-h5cloud2023}
ICESAT-2 HackWeek, H. C. (2023). h5cloud: Tools for cloud-based analysis
of HDF5 data (Version v1.0.0). Retrieved from
\url{https://github.com/ICESAT-2HackWeek/h5cloud}

\bibitem[\citeproctext]{ref-Mozilla-latency-2024}
MDN, M. (2024, May). Understanding latency. Retrieved from
\url{https://developer.mozilla.org/en-US/docs/Web/Performance/Understanding_latency}

\bibitem[\citeproctext]{ref-NEUMANN2019111325}
Neumann, T. A., Martino, A. J., Markus, T., Bae, S., Bock, M. R.,
Brenner, A. C., et al. (2019). The ice, cloud, and land elevation
satellite -- 2 mission: A global geolocated photon product derived from
the advanced topographic laser altimeter system. \emph{Remote Sensing of
Environment}, \emph{233}, 111325.
https://doi.org/\url{https://doi.org/10.1016/j.rse.2019.111325}

\bibitem[\citeproctext]{ref-scott-2020}
Scott, C. (2020). Numbers every programmer should know. Retrieved from
\url{https://colin-scott.github.io/personal_website/research/interactive_latency.html}

\bibitem[\citeproctext]{ref-The_HDF_Group_Hierarchical_Data_Format}
The HDF Group. (n.d.). {Hierarchical Data Format, version 5}. Retrieved
from \url{https://github.com/HDFGroup/hdf5}

\end{CSLReferences}

\end{document}